Project: Cyberbullying Detection:
Dataset One:

Step 1:
The goal is to build a predictive model that can classify textual content (such as social media posts, comments, and messages) as either cyberbullying or non-cyberbullying. Cyberbullying can take many forms, including aggression, attack, toxicity, racism, sexism, and other harmful behaviors in online platforms. By building an automated system, we aim to identify harmful or offensive behavior in text and reduce the spread of such content in online communities
```{r}
options(repos = c(CRAN = "https://cloud.r-project.org/"))

# Install all the necessary packages
install.packages("janeaustenr")
install.packages("wordcloud")
install.packages("tm")
install.packages("text2vec")
install.packages("dplyr")
install.packages("tidyr")
install.packages("tidytext")
install.packages("ggplot2")
install.packages("stopwords")
install.packages("lme4")
install.packages("nlme")
install.packages("caret")
install.packages("SnowballC")
install.packages("ROCR")  # Ensure ROCR is installed

```
Step 2:


```{r}
library(janeaustenr)
library(wordcloud)
library(tm)
library(text2vec)
library(dplyr)
library(tidyr)
library(tidytext)
library(ggplot2)
library(stopwords)
library(lme4)
library(nlme)
library(caret)
library(SnowballC)
library(ROCR)


```

```{r}
aggressive=read.csv("C:\\Users\\aabua\\Desktop\\apd\\archive\\aggression_parsed_dataset.csv")
head(aggressive)
```


```{r}
attack=read.csv("C:\\Users\\aabua\\Desktop\\apd\\archive\\attack_parsed_dataset.csv")
head(attack)
```

```{r}
kaggle=read.csv("C:\\Users\\aabua\\Desktop\\apd\\archive\\kaggle_parsed_dataset.csv")
head(kaggle)
```

```{r}
toxicity=read.csv("C:\\Users\\aabua\\Desktop\\apd\\archive\\toxicity_parsed_dataset.csv")
head(toxicity)
```

```{r}
twitter_parsed=read.csv("C:\\Users\\aabua\\Desktop\\apd\\archive\\twitter_parsed_dataset.csv")
head(twitter_parsed)
```

```{r}
twitter_racism=read.csv("C:\\Users\\aabua\\Desktop\\apd\\archive\\twitter_racism_parsed_dataset.csv")
head(twitter_racism)
```

```{r}
twitter_sexism=read.csv("C:\\Users\\aabua\\Desktop\\apd\\archive\\twitter_sexism_parsed_dataset.csv")
head(twitter_sexism)

```

```{r}
youtube_parsed=read.csv("C:\\Users\\aabua\\Desktop\\apd\\archive\\youtube_parsed_dataset.csv")
head(youtube_parsed)
```


step 3:


```{r}

# Load necessary libraries
library(dplyr)  # For %>% operator and other dplyr functions
library(tidyr)  # For drop_na() function

library(stringr)

# Select relevant columns from each dataset and add appropriate annotations
aggressive <- aggressive %>% 
  select(Text, oh_label)

attack <- attack %>% 
  select(Text, oh_label)

kaggle <- kaggle %>% 
  select(Text, oh_label)

toxicity <- toxicity %>% 
  select(Text, oh_label) 

twitter_parsed <- twitter_parsed %>% 
  select(Text, oh_label)

twitter_racism <- twitter_racism %>% 
  select(Text, oh_label)

twitter_sexism <- twitter_sexism %>% 
  select(Text, oh_label)

youtube_parsed <- youtube_parsed %>% 
  select(Text, oh_label) 

# Merge datasets
merged_data <- bind_rows(
  aggressive,
  attack,
  kaggle,
  toxicity,
  twitter_parsed %>% select(Text, oh_label),
  twitter_racism %>% select(Text, oh_label),
  twitter_sexism %>% select(Text, oh_label),
  youtube_parsed %>% select(Text, oh_label)
)

# Add a unique index to the dataset
cyberbullying_classification_data <- merged_data %>%
  mutate(Index = row_number()) %>%
  select(Index, everything())  # Reorder to place Index as the first column

# Step 1: Remove duplicates
cyberbullying_classification_data <- cyberbullying_classification_data %>%
  distinct(Text, .keep_all = TRUE)  # Remove duplicate rows based on 'Text' column

# Check for duplicates
duplicated_rows <- sum(duplicated(cyberbullying_classification_data$Text))  # Number of duplicate rows
cat("Number of duplicate rows removed:", duplicated_rows, "\n")

# Step 2: Handle missing values

# Check how many missing values there are in the 'Text' column
missing_values_count <- sum(is.na(cyberbullying_classification_data$Text))
cat("Number of missing values in 'Text' column:", missing_values_count, "\n")

# Remove rows with missing 'Text' and 'oh_label' columns
cyberbullying_classification_data <- cyberbullying_classification_data %>%
  drop_na(Text, oh_label)  # Removing rows with missing 'Text' and 'oh_label'

# Step 3: Clean text data - Remove unnecessary characters (HTML tags, URLs, punctuation)
# Step 3: Clean text data - Remove unnecessary characters (HTML tags, URLs, punctuation, special symbols)
cyberbullying_classification_data <- cyberbullying_classification_data %>%
  mutate(Text = str_remove_all(Text, "<.*?>")) %>%         # Remove HTML tags
  mutate(Text = str_remove_all(Text, "http\\S+|www\\S+")) %>%  # Remove URLs
  mutate(Text = str_remove_all(Text, "[[:punct:]]")) %>%    # Remove punctuation
  mutate(Text = str_remove_all(Text, "`")) %>%              # Remove backticks (`)
  mutate(Text = str_replace_all(Text, "[^[:alnum:]\\s]", ""))

# Remove numbers (if needed)
cyberbullying_classification_data <- cyberbullying_classification_data %>%
  mutate(Text = str_replace_all(Text, "\\d+", ""))  # Remove numbers




# Step 4: Lowercase conversion - Convert all text to lowercase
cyberbullying_classification_data <- cyberbullying_classification_data %>%
  mutate(Text = str_to_lower(Text))  # Convert to lowercase

# Final cleaned dataset - Preview

head(cyberbullying_classification_data)

# Write the cleaned dataset to a CSV file
write.csv(cyberbullying_classification_data, "merged_cleaned_dataset_with_tokens.csv", row.names = FALSE)

```
```{r}
summary(cyberbullying_classification_data)
# Inspect the structure
str(cyberbullying_classification_data)

table(cyberbullying_classification_data$oh_label)


```
```{r}
# Load necessary libraries for visualization

# Add a new column for text length
cyberbullying_classification_data$Text_length <- nchar(cyberbullying_classification_data$Text)

# Plot text length distribution for both classes
ggplot(cyberbullying_classification_data, aes(x = Text_length, fill = as.factor(oh_label))) +
  geom_histogram(binwidth = 50, position = "dodge") +
  labs(title = "Text Length Distribution by Class", x = "Text Length", y = "Frequency") +
  scale_fill_manual(values = c("blue", "red"), labels = c("Non-Cyberbullying", "Cyberbullying")) +
  theme_minimal()


```

```{r}
# Plot the class distribution
ggplot(cyberbullying_classification_data, aes(x = as.factor(oh_label), fill = as.factor(oh_label))) +
  geom_bar() +
  labs(title = "Class Distribution (Cyberbullying vs Non-Cyberbullying)", 
       x = "Class (0 = Non-Cyberbullying, 1 = Cyberbullying)", 
       y = "Count") +
  scale_fill_manual(values = c("blue", "red")) +
  theme_minimal()

```

```{r}
# Load the 'dplyr' and 'tidytext' libraries

# Tokenize the text and get word frequency counts
word_freq <- cyberbullying_classification_data %>%
  unnest_tokens(word, Text) %>%
  anti_join(stop_words) %>%  # Remove common stop words
  group_by(oh_label, word) %>%
  tally(sort = TRUE) %>%
  ungroup()

# Get top 10 most frequent words for each class
top_words <- word_freq %>%
  group_by(oh_label) %>%
  top_n(10, n) %>%
  arrange(oh_label, desc(n))

# Plot the top 10 frequent words by class
ggplot(top_words, aes(x = reorder(word, n), y = n, fill = as.factor(oh_label))) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 10 Most Frequent Words by Class", x = "Word", y = "Frequency") +
  scale_fill_manual(values = c("blue", "red"), labels = c("Non-Cyberbullying", "Cyberbullying")) +
  theme_minimal()


```
step 5
```{r}
library(tm)
library(tm)

# Sample 5% of the dataset to reduce memory usage
sample_size <- floor(0.01 * nrow(cyberbullying_classification_data))  # Calculate 5% of rows
sample_indices <- sample(seq_len(nrow(cyberbullying_classification_data)), size = sample_size)  # Random sampling
 
sampled_data <- cyberbullying_classification_data[sample_indices, ]  # Get the subset

# Create a corpus from the sampled text data
corpus <- Corpus(VectorSource(sampled_data$Text))

# Create Document-Term Matrix (DTM)
dtm <- DocumentTermMatrix(corpus)

# Convert DTM to a dataframe
dtm_df <- as.data.frame(as.matrix(dtm))

# Add labels to the dataset
dtm_df$oh_label <- sampled_data$oh_label

# View a portion of the DTM dataframe
head(dtm_df)





```

step 6-  
```{r}

# Load necessary libraries
library(caret)

# Set seed for reproducibility
set.seed(123)

# Split data into training (70%) and test (30%) sets
trainIndex <- createDataPartition(dtm_df$oh_label, p = 0.7, list = FALSE)
trainData <- dtm_df[trainIndex, ]
testData <- dtm_df[-trainIndex, ]

```


```{r}
# Load necessary library
library(glmnet)

# Prepare training data (excluding the label column)
x_train <- as.matrix(trainData[, -ncol(trainData)])  # Exclude label column
y_train <- trainData$oh_label  # Dependent variable (label)

# Step 1: Fit Ridge Regression Model (alpha = 0 for Ridge)
ridge_model <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0)

# Print the Ridge model output
print(ridge_model)

# Step 2: Make predictions for Ridge model
ridge_predictions <- predict(ridge_model, newx = as.matrix(testData[, -ncol(testData)]), type = "response")

# Convert probabilities to class labels (0 or 1)
ridge_pred_class <- ifelse(ridge_predictions > 0.5, 1, 0)  # If probability > 0.5, assign class 1, otherwise class 0

```


```{r}
# Load necessary library
library(glmnet)

# Prepare training data (excluding the label column)
x_train <- as.matrix(trainData[, -ncol(trainData)])  # Exclude label column
y_train <- trainData$oh_label  # Dependent variable (label)

# Step 1: Fit Lasso Regression Model (alpha = 1 for Lasso)
lasso_model <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 1)

# Print the Lasso model output
print(lasso_model)

# Step 2: Make predictions for Lasso model
lasso_predictions <- predict(lasso_model, newx = as.matrix(testData[, -ncol(testData)]), type = "response")

# Convert probabilities to class labels (0 or 1)
lasso_pred_class <- ifelse(lasso_predictions > 0.5, 1, 0)  # If probability > 0.5, assign class 1, otherwise class 0
```
step 7


```{r}
# Load necessary libraries for performance evaluation
library(caret)
install.packages("ROCR")  # Ensure ROCR is installed for ROC/AUC analysis
library(ROCR)

# Step 1: Confusion Matrix for Ridge Model
ridge_confusion <- confusionMatrix(as.factor(ridge_pred_class), as.factor(testData$oh_label))
cat("Ridge Model Performance:\n")
print(ridge_confusion)

# Step 2: Confusion Matrix for Lasso Model
lasso_confusion <- confusionMatrix(as.factor(lasso_pred_class), as.factor(testData$oh_label))
cat("Lasso Model Performance:\n")
print(lasso_confusion)

# Step 3: Calculate Accuracy, Precision, Recall, and F1-Score for Ridge Model
ridge_accuracy <- ridge_confusion$overall['Accuracy']
ridge_precision <- ridge_confusion$byClass['Pos Pred Value']
ridge_recall <- ridge_confusion$byClass['Sensitivity']
ridge_f1 <- 2 * ((ridge_precision * ridge_recall) / (ridge_precision + ridge_recall))

cat("\nRidge Model Metrics:\n")
cat("Accuracy:", ridge_accuracy, "\n")
cat("Precision:", ridge_precision, "\n")
cat("Recall:", ridge_recall, "\n")
cat("F1-Score:", ridge_f1, "\n")

# Step 4: Calculate Accuracy, Precision, Recall, and F1-Score for Lasso Model
lasso_accuracy <- lasso_confusion$overall['Accuracy']
lasso_precision <- lasso_confusion$byClass['Pos Pred Value']
lasso_recall <- lasso_confusion$byClass['Sensitivity']
lasso_f1 <- 2 * ((lasso_precision * lasso_recall) / (lasso_precision + lasso_recall))

cat("\nLasso Model Metrics:\n")
cat("Accuracy:", lasso_accuracy, "\n")
cat("Precision:", lasso_precision, "\n")
cat("Recall:", lasso_recall, "\n")
cat("F1-Score:", lasso_f1, "\n")

# Step 5: ROC Curve for Ridge Model
ridge_pred_prob <- ridge_predictions[,1]  # Probabilities for class 1
roc_ridge <- prediction(ridge_pred_prob, testData$oh_label)
roc_ridge_perf <- performance(roc_ridge, measure = "tpr", x.measure = "fpr")

# Plot ROC curve for Ridge
plot(roc_ridge_perf, main = "ROC Curve for Ridge Model", col = "blue", lwd = 2)

# Step 6: ROC Curve for Lasso Model
lasso_pred_prob <- lasso_predictions[,1]  # Probabilities for class 1
roc_lasso <- prediction(lasso_pred_prob, testData$oh_label)
roc_lasso_perf <- performance(roc_lasso, measure = "tpr", x.measure = "fpr")

# Plot ROC curve for Lasso
plot(roc_lasso_perf, main = "ROC Curve for Lasso Model", col = "red", lwd = 2, add = TRUE)

# Step 7: AUC for Ridge and Lasso Models
roc_ridge_auc <- performance(roc_ridge, measure = "auc")@y.values[[1]]
roc_lasso_auc <- performance(roc_lasso, measure = "auc")@y.values[[1]]

cat("\nAUC for Ridge Model:", roc_ridge_auc, "\n")
cat("AUC for Lasso Model:", roc_lasso_auc, "\n")

```


step 8
```{r}
# Load necessary library
install.packages("doParallel")  # Run this only once
library(caret)
library(doParallel)
# Set up a smaller grid for Ridge Regression
ridge_grid <- expand.grid(alpha = 0,  # Ridge: alpha = 0
                          lambda = seq(0.001, 0.01, length = 3))

# Set up a smaller grid for Lasso Regression
lasso_grid <- expand.grid(alpha = 1,  # Lasso: alpha = 1
                          lambda = seq(0.001, 0.01, length = 3))

# Set up cross-validation (5-fold)
train_control <- trainControl(method = "cv", number = 5)

# Register parallel backend
numCores <- detectCores() - 1  # Use one less core
cl <- makeCluster(numCores)
registerDoParallel(cl)

# Ridge Regression Grid Search
ridge_grid_search <- train(oh_label ~ ., 
                           data = trainData, 
                           method = "glmnet", 
                           trControl = train_control, 
                           tuneGrid = ridge_grid)

# Lasso Regression Grid Search
lasso_grid_search <- train(oh_label ~ ., 
                           data = trainData, 
                           method = "glmnet", 
                           trControl = train_control, 
                           tuneGrid = lasso_grid)

# Stop parallel backend properly
stopCluster(cl)
registerDoSEQ()  # Reset backend to sequential mode

# Print best parameters and model performance
print(ridge_grid_search)
print(lasso_grid_search)
```
Dataset 2
step1
The objective of this project is to detect cyberbullying in social media comments by classifying each comment into categories: Normal, Offensive, and Hate Speech. Additionally, comments will be categorized based on specific Demographic Groups, such as Race, Religion, Gender, Sexual Orientation, and Miscellaneous. The goal is to identify bullying comments and categorize the specific demographic groups being targeted. This involves building a multi-class classification model to predict the bullying status and, if applicable, the targeted demographic group
step2
```{r}
hateXplain=read.csv("C:\\Users\\aabua\\Desktop\\apd\\work2\\final_hateXplain.csv")
head(hateXplain)
```



step 3
```{r}
library(dplyr)
library(stopwords)

library(stringr)

# Load the hateXplain dataset (make sure the path is correct)
hateXplain <- read.csv("C:\\Users\\aabua\\Desktop\\apd\\work2\\final_hateXplain.csv")

# Step 1: View the first few rows of the dataset to check structure
head(hateXplain)

# Step 2: Ensure the 'comment' column is in character format (if not already)
hateXplain$comment <- as.character(hateXplain$comment)

# Step 3: Remove duplicates based on the 'comment' column
hateXplain <- hateXplain %>%
  distinct(comment, .keep_all = TRUE)  # Remove duplicate rows based on 'comment' column

# Step 4: Handle missing values
missing_values_count <- sum(is.na(hateXplain$comment))
cat("Number of missing values in 'comment' column:", missing_values_count, "\n")
hateXplain <- hateXplain %>%
  drop_na(comment)  # Remove rows with missing 'comment' column

# Step 5: Clean text data - Remove unnecessary characters (HTML tags, URLs, punctuation, special symbols)
hateXplain <- hateXplain %>%
  mutate(comment = str_remove_all(comment, "<.*?>")) %>%            # Remove HTML tags (this should work)
  mutate(comment = str_remove_all(comment, "http[s]?://\\S+|www\\S+")) %>%  # Improved URL matching (handles https)
  mutate(comment = str_remove_all(comment, "[[:punct:]]")) %>%       # Remove punctuation
  mutate(comment = str_remove_all(comment, "`")) %>%                 # Remove backticks (`)
  mutate(comment = str_replace_all(comment, "[^[:alnum:]\\s]", ""))  # Remove special symbols

# Step 6: Remove numbers (if needed)
hateXplain <- hateXplain %>%
  mutate(comment = str_replace_all(comment, "\\d+", ""))  # Remove numbers
# Step 6: Remove specific words like "th" and "o"
hateXplain <- hateXplain %>%
  mutate(comment = str_replace_all(comment, "\\b(th|o)\\b", ""))  # Remove standalone words "th" and "o"

# Step 7: Lowercase conversion - Convert all text to lowercase
hateXplain <- hateXplain %>%
  mutate(comment = str_to_lower(comment))  # Convert to lowercase
# Step 8: Remove stopwords
# Using the 'tm' package to get a list of common stopwords in English
stop_words <- stopwords("en")
hateXplain <- hateXplain %>%
  mutate(comment = str_remove_all(comment, paste("\\b", stop_words, "\\b", collapse = "|")))  # Remove stopwords
hateXplain <- hateXplain %>%
  mutate(comment = str_remove_all(comment, "[\U0001F600-\U0001F64F]"))  # Emojis in the Unicode range 1F600-1F64F (Emoticons)
# Step 10: Remove one-letter words
hateXplain <- hateXplain %>%
  mutate(comment = str_remove_all(comment, "\\b[a-zA-Z]\\b"))  # Remove one-letter words



# Final check on the cleaned data
head(hateXplain)

# Step 8: Save the cleaned dataset to a CSV file
write.csv(hateXplain, "C:\\Users\\aabua\\Desktop\\apd\\work2\\cleaned_hateXplain.csv", row.names = FALSE)

cat("Cleaning complete. The cleaned dataset has been saved as 'cleaned_hateXplain.csv'.\n")

```
```{r}
summary(cleaned_hateXplain)
# Inspect the structure

table(cleaned_hateXplain$label)


```

```{r}
# Load necessary libraries
library(ggplot2)
library(readr)
cleaned_hateXplain <- read.csv("C:/Users/aabua/Desktop/apd/work2/cleaned_hateXplain.csv")


# Check unique values of the 'label' column to verify the exact levels
unique(cleaned_hateXplain$label)

# Calculate the frequency of each label (Hatespeech, Offensive, Normal)
label_counts <- table(cleaned_hateXplain$label)

# Convert to a data frame for ggplot2
label_df <- as.data.frame(label_counts)

# Calculate percentages for each category
label_df$percent <- (label_df$Freq / sum(label_df$Freq)) * 100

# Create pie chart with percentages
pie_chart <- ggplot(data = label_df, aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Distribution of  Labels") +
  scale_fill_manual(values = c("hatespeech" = "red", "offensive" = "orange", "normal" = "green")) +
  geom_text(aes(label = paste0(round(percent, 1), "%")), position = position_stack(vjust = 0.5), size = 5)

# Print the pie chart
print(pie_chart)

```

```{r}
# Check unique values in 'label' column
unique(cleaned_hateXplain$label)

# Histogram for Hate Speech vs Race
ggplot(cleaned_hateXplain, aes(x = Race, fill = label)) +
  geom_bar(position = "dodge") +
  labs(title = "Label vs Race", x = "Race", y = "Count") +
  scale_fill_manual(values = c("hatespeech" = "red", "offensive" = "orange", "normal" = "green"))

```

```{r}
# Check unique values in 'label' column
unique(cleaned_hateXplain$label)

# Histogram for Hate Speech vs Race
ggplot(cleaned_hateXplain, aes(x = Religion, fill = label)) +
  geom_bar(position = "dodge") +
  labs(title = "Label vs Religion", x = "Religion", y = "Count") +
  scale_fill_manual(values = c("hatespeech" = "red", "offensive" = "orange", "normal" = "green"))

```

```{r}
# Check unique values in 'label' column
unique(cleaned_hateXplain$label)

# Histogram for Hate Speech vs Race
ggplot(cleaned_hateXplain, aes(x = Gender, fill = label)) +
  geom_bar(position = "dodge") +
  labs(title = "Label vs Gender", x = "Gender", y = "Count") +
  scale_fill_manual(values = c("hatespeech" = "red", "offensive" = "orange", "normal" = "green"))

```

```{r}
# Check unique values in 'label' column
unique(cleaned_hateXplain$label)

# Histogram for Hate Speech vs Race
ggplot(cleaned_hateXplain, aes(x = Sexual.Orientation, fill = label)) +
  geom_bar(position = "dodge") +
  labs(title = "Label vs Sexual Orientation", x = "Sexual Orientation", y = "Count") +
  scale_fill_manual(values = c("hatespeech" = "red", "offensive" = "orange", "normal" = "green"))

```

```{r}
# Check unique values in 'label' column
unique(cleaned_hateXplain$label)

# Histogram for Hate Speech vs Race
ggplot(cleaned_hateXplain, aes(x = Miscellaneous, fill = label)) +
  geom_bar(position = "dodge") +
  labs(title = "Label vs Miscellaneous", x = "Miscellaneous", y = "Count") +
  scale_fill_manual(values = c("hatespeech" = "red", "offensive" = "orange", "normal" = "green"))

```

```{r}
install.packages("tidyverse")

```
```{r}
# Load necessary libraries
library(ggplot2)
library(tidytext)
library(tidyverse)

library(wordcloud)

# Load the dataset
cleaned_hateXplain <- read.csv("C:/Users/aabua/Desktop/apd/work2/cleaned_hateXplain.csv")

# Ensure case consistency in the Text column (if it's in the 'comment' column, change as needed)
cleaned_hateXplain$comment <- tolower(cleaned_hateXplain$comment)  # Make text lowercase

# Tokenize the text data to get word counts by label
word_freq <- cleaned_hateXplain %>%
  unnest_tokens(word, comment) %>%  # 'comment' is the text column, adjust if your column name is different
  count(word, label, sort = TRUE)    # Count words by 'label' (e.g., Hatespeech, Normal, Offensive)

# Generate Word Cloud with Colors
wordcloud(words = word_freq$word, 
          freq = word_freq$n, 
          min.freq = 10, 
          max.words = 100, 
          random.order = FALSE, 
          colors = c("blue", "red", "green")[as.factor(word_freq$label)])

```
step 5

```{r}
# Install necessary packages only if they are not already installed
required_packages <- c("tm", "text2vec", "dplyr", "caret", "class", "ggplot2", "randomForest", "caTools")

installed_packages <- installed.packages()
missing_packages <- setdiff(required_packages, rownames(installed_packages))

if(length(missing_packages) > 0) {
  install.packages(missing_packages)
}

# Load the libraries
library(tm)
library(text2vec)
library(dplyr)
library(caret)
library(class)
library(ggplot2)
library(randomForest)
library(caTools)

# Step 1: Load and Sample Data (5% of the dataset)
# Load the cleaned dataset (assuming it is named cleaned_hateXplain)
cleaned_hateXplain <- read.csv("C:/Users/aabua/Desktop/apd/work2/cleaned_hateXplain.csv")  # Replace with actual path if necessary

# Sample 1% of the dataset to reduce memory usage (Adjust this to 5% if required)
sample_size <- floor(0.01 * nrow(cleaned_hateXplain))  # Calculate 5% of rows
sample_indices <- sample(seq_len(nrow(cleaned_hateXplain)), size = sample_size)  # Random sampling

sampled_data <- cleaned_hateXplain[sample_indices, ]  # Subset the data to sampled portion

# Step 2: Preprocess the 'comment' column (text data)
comments <- sampled_data$comment  # Assuming the 'comment' column exists

# Convert text to lower case
comments <- tolower(comments)

# Remove special characters and digits
comments <- gsub("[^a-zA-Z\\s]", "", comments)

# Remove stopwords (common words that do not add much value)
comments <- removeWords(comments, stopwords("en"))

# Optional: Perform stemming (reduces words to their root form)
comments <- stemDocument(comments)

# Step 3: Convert the cleaned text into a Document-Term Matrix (DTM) using TF-IDF
dtm <- DocumentTermMatrix(Corpus(VectorSource(comments)))

# Convert DTM into a matrix
dtm_matrix <- as.matrix(dtm)

# View the resulting TF-IDF matrix (first few rows)
head(dtm_matrix)

# Step 4: Combine TF-IDF features with other non-text features (like Race, Religion, etc.)
# Assuming your other features are in columns such as 'Race', 'Religion', 'Gender', etc.
# Extract these non-text features from the cleaned dataset
non_text_features <- sampled_data %>%
  select(Race, Religion, Gender, Sexual.Orientation, Miscellaneous)  # Adjust column names accordingly

# Combine the TF-IDF matrix with these non-text features
X_text <- as.data.frame(dtm_matrix)  # Convert TF-IDF matrix to data frame
X_all <- cbind(X_text, non_text_features)  # Combine TF-IDF with other features

# View the combined dataset
head(X_all)

# Step 5: Perform Feature Selection using Chi-Square Test
# Perform chi-square test to select relevant features based on p-value
y <- sampled_data$label  # Assuming 'label' is the target column

# Chi-square test for each feature
chi_sq_results_all <- apply(X_all, 2, function(x) {
  chisq.test(table(x, y))$p.value
})

# View the Chi-Square p-values for each feature
print(chi_sq_results_all)

# Select features with p-value < 0.05 (statistically significant features)
selected_features_all <- names(chi_sq_results_all[chi_sq_results_all < 0.05])

# Filter the dataset to include only the selected features
X_selected_all <- X_all[, selected_features_all]

# View the selected features
print(selected_features_all)

```
step 6
```{r}
# Split the data into training and testing sets
set.seed(123)
split <- sample.split(y, SplitRatio = 0.7)

# Create training and testing sets
train_data <- subset(X_selected_all, split == TRUE)
test_data <- subset(X_selected_all, split == FALSE)

# Create the labels for training and testing sets
train_labels <- y[split == TRUE]
test_labels <- y[split == FALSE]

# Check the first few elements of train_labels and test_labels
head(train_labels)
head(test_labels)
# Check if there are any NULL or NA values in train_labels and test_labels
sum(is.null(train_labels))  # Should return 0
sum(is.null(test_labels))   # Should return 0
sum(is.na(train_labels))    # Should return 0
sum(is.na(test_labels))     # Should return 0
# Ensure both train_labels and test_labels have the same factor levels
train_labels <- factor(train_labels, levels = c("normal", "hatespeech", "offensive"))
test_labels <- factor(test_labels, levels = c("normal", "hatespeech", "offensive"))

```

```{r}
#install.packages("nnet")  # Install the nnet package if you don't have it already
library(nnet)  # Load the nnet package

# Train the multinomial logistic regression model
multinom_model <- multinom(train_labels ~ ., data = train_data)

# Print the model summary to understand the coefficients
summary(multinom_model)
# Make predictions on the test set (returns class labels)
predicted_class_multinom <- predict(multinom_model, test_data)

# View the first few predictions
head(predicted_class_multinom)

```



```{r}
# Ensure target labels are factors with all levels
all_levels <- levels(train_labels)  # Get all levels from the training set
test_labels <- factor(test_labels, levels = all_levels)  # Ensure test set has all levels

# Encode the factor labels as numeric values
train_labels_numeric <- as.numeric(train_labels)
test_labels_numeric <- as.numeric(test_labels)

# Linear Model: Fit the model with numeric labels
linear_model <- lm(train_labels_numeric ~ ., data = train_data)

# Make predictions on the test set
predictions_linear <- predict(linear_model, test_data)

# Round predictions to the nearest integer (since labels are categorical)
predictions_linear <- round(predictions_linear)

# Convert numeric predictions back to factor
predictions_linear <- factor(predictions_linear, levels = c(1, 2, 3))  # Assuming 1, 2, 3 represent the classes

# Evaluate performance
confusion_matrix_linear <- table(predictions_linear, test_labels)
print(confusion_matrix_linear)

```

step 7

```{r}
# Confusion Matrix for multinomial logistic regression model
conf_matrix_log <- table(predicted_class_multinom, test_labels)
print(conf_matrix_log)

# Accuracy: Overall percentage of correct predictions
accuracy_log <- sum(diag(conf_matrix_log)) / sum(conf_matrix_log)
cat("Accuracy: ", accuracy_log, "\n")

# Precision, Recall, F1 for each label (Hate Speech, Offensive, Normal)
precision_log <- diag(conf_matrix_log) / rowSums(conf_matrix_log)
recall_log <- diag(conf_matrix_log) / colSums(conf_matrix_log)
f1_log <- 2 * (precision_log * recall_log) / (precision_log + recall_log)

cat("Precision (Hate Speech): ", precision_log["hatespeech"], "\n")
cat("Recall (Hate Speech): ", recall_log["hatespeech"], "\n")
cat("F1-Score (Hate Speech): ", f1_log["hatespeech"], "\n")

cat("Precision (Offensive): ", precision_log["offensive"], "\n")
cat("Recall (Offensive): ", recall_log["offensive"], "\n")
cat("F1-Score (Offensive): ", f1_log["offensive"], "\n")

cat("Precision (Normal): ", precision_log["normal"], "\n")
cat("Recall (Normal): ", recall_log["normal"], "\n")
cat("F1-Score (Normal): ", f1_log["normal"], "\n")

# Specificity (True Negative Rate) for each class (generalized)
specificity_log <- (rowSums(conf_matrix_log) - diag(conf_matrix_log)) / (sum(conf_matrix_log) - rowSums(conf_matrix_log))
cat("Specificity (Hate Speech): ", specificity_log["hatespeech"], "\n")
cat("Specificity (Offensive): ", specificity_log["offensive"], "\n")
cat("Specificity (Normal): ", specificity_log["normal"], "\n")



```

```{r}
# Ensure train_labels and test_labels are factors
train_labels <- factor(train_labels, levels = c("hatespeech", "offensive", "normal"))
test_labels <- factor(test_labels, levels = c("hatespeech", "offensive", "normal"))

# Encode the factor labels as numeric values
train_labels_numeric <- as.numeric(train_labels)
test_labels_numeric <- as.numeric(test_labels)

# Linear Model: Fit the model with numeric labels
linear_model <- lm(train_labels_numeric ~ ., data = train_data)

# Make predictions on the test set
predictions_linear <- predict(linear_model, test_data)

# Round predictions to the nearest integer (since labels are categorical)
predictions_linear <- round(predictions_linear)

# Convert numeric predictions back to factor (assuming 1 = "hatespeech", 2 = "offensive", 3 = "normal")
predictions_linear <- factor(predictions_linear, levels = c(1, 2, 3))  

# Confusion Matrix for Linear Regression model
conf_matrix_linear <- table(predictions_linear, test_labels)
print(conf_matrix_linear)

# Accuracy: Overall percentage of correct predictions
accuracy_linear <- sum(diag(conf_matrix_linear)) / sum(conf_matrix_linear)
cat("Accuracy: ", accuracy_linear, "\n")

# Precision, Recall, F1 for each class (Hate Speech, Offensive, Normal)
precision_linear <- diag(conf_matrix_linear) / rowSums(conf_matrix_linear)
recall_linear <- diag(conf_matrix_linear) / colSums(conf_matrix_linear)
f1_linear <- 2 * (precision_linear * recall_linear) / (precision_linear + recall_linear)

cat("Precision (Hate Speech): ", precision_linear["1"], "\n")
cat("Recall (Hate Speech): ", recall_linear["1"], "\n")
cat("F1-Score (Hate Speech): ", f1_linear["1"], "\n")

cat("Precision (Offensive): ", precision_linear["2"], "\n")
cat("Recall (Offensive): ", recall_linear["2"], "\n")
cat("F1-Score (Offensive): ", f1_linear["2"], "\n")

cat("Precision (Normal): ", precision_linear["3"], "\n")
cat("Recall (Normal): ", recall_linear["3"], "\n")
cat("F1-Score (Normal): ", f1_linear["3"], "\n")

# Specificity (True Negative Rate) for each class (generalized)
specificity_linear <- (rowSums(conf_matrix_linear) - diag(conf_matrix_linear)) / (sum(conf_matrix_linear) - rowSums(conf_matrix_linear))
cat("Specificity (Hate Speech): ", specificity_linear["1"], "\n")
cat("Specificity (Offensive): ", specificity_linear["2"], "\n")
cat("Specificity (Normal): ", specificity_linear["3"], "\n")

# ROC Curve and AUC (Area Under the Curve) for Linear Model
library(pROC)

# ROC Curve Calculation (ensure the prediction is numeric for multiclass ROC)
roc_curve_linear <- multiclass.roc(test_labels, as.numeric(predictions_linear))
auc_linear <- auc(roc_curve_linear)
cat("AUC for Linear Regression: ", auc_linear, "\n")


```


step 8


